---
title: "Unveiling the Ticket Price Dynamics: A Deep Dive into Tomaticket's Show Prices"
author: "Andrea Baños and Daniel Pérez"
date: "2024-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the main libraries

```{r, include=TRUE, warning=FALSE}
library(rvest)
library(xml2)
library(tidyverse)
library(leaflet)
library(plotly)

```



```{r, include=TRUE, warning=FALSE}
url <- "https://www.tomaticket.es/index.php?accion=search&buscador=&IdLugar=0&IdTag=4"

tomaticket <- read_html(url) |> xml_child()


```

## Navigating through the extracted data - Generating the variables

```{r}
show <-
  tomaticket %>%
  xml_find_all("//div[@id='list-search-event' and @class='search-events-normal']//a") %>%
  xml_attr(attr = "href")
show

urls_clean <- show[!grepl("^https:", show)]
show_links <- paste0("https://www.tomaticket.es/", urls_clean)
show_links


theaters <- c()

for (link in show_links) {
  event_page <- read_html(link)
  theater <- event_page %>%
    html_node(xpath = "//div[@id='EventHeadContenido' and @class='white-space']//p[@class='nombre-recinto']//a") %>%
    html_attr("href")
  theaters <- c(theaters, theater)
}

```

```{r}
coordenadas <- c()
for (link in show_links) {
  event_page <- read_html(link)
  
  localizacion <- event_page %>%
    html_node(xpath = "//a[contains(@class, 'sidebar-links') and contains(@href, 'maps.google.com')]/@href") %>%
    html_text()
  coordenadas <- c(coordenadas,localizacion)
  
  
}
df_coordenadas <- data.frame(coordenada = coordenadas)
df_coordenadas$coordenada <- sub(".*/?q=(.*)$", "\\1", df_coordenadas$coordenada)

df_1 <- as_tibble(df_coordenadas)
df_1<- separate(df_1, coordenada, into = c("Latitud", "Longitud"), sep = ",")
df_1$id <- c(1:100)
```

```{r}
only_prices <- tomaticket %>%
  xml_find_all("//p[@class='bottommargin-precio' and @itemprop='offers' and @itemscope='' and @itemtype='https://schema.org/AggregateOffer']//meta[@itemprop='price']")
only_prices

datetime <- tomaticket %>%
  xml_find_all("//p[@class='bottommargin-precio' and @itemprop='offers' and @itemscope='' and @itemtype='https://schema.org/AggregateOffer']//time")
datetime

location <- tomaticket %>%
  xml_find_all("//p[@class='lugar' and @itemprop='location' and @itemscope='' and @itemtype='https://schema.org/Place']//meta")
location

```

## Data Cleaning and Final Data Frame Creation

```{r}
df_theaters <- data.frame(theater = theaters)
df_shows <- data.frame(show_links)

clean_dates <- sub("^[^=]+=[^=]+=(.*)$", "\\1", datetime)
clean_dates <- sub(".*\"(\\d{4}-\\d{2}-\\d{2}).*", "\\1", clean_dates)
dates <- as.Date(clean_dates)

df_date <- data.frame(dates)
df_date$id <- c(1:100)

df_theaters$theater_name <- sub(".*/recintos/(.*)$", "\\1", df_theaters$theater)
df_theaters <- df_theaters %>% 
  select(-theater)

df_theaters$id <- c(1:100)
final_df <- left_join(df_theaters, df_date, by = "id")

df_shows$show <- sub(".*/entradas-(.*)", "\\1", df_shows$show_links)
df_shows <- df_shows %>% 
  select(-show_links)
df_shows$id <- c(1:100)
final_df <- left_join(final_df, df_shows, by = "id")

location_address <- grep('address', location, value = TRUE)
clean_location <- sub('.*="([^"]*)".*', '\\1', location_address)
df_location <- data.frame(clean_location)
df_location$id <- c(1:100)
final_df <- left_join(final_df, df_location, by = "id")

clean_prices <- sub('.*="([^"]*)".*', '\\1', only_prices)
df_prices <- data.frame(clean_prices)
df_prices$id <- c(1:100)
df_prices$clean_prices <- gsub(",", ".", df_prices$clean_prices)

df_prices$clean_prices <- as.numeric(df_prices$clean_prices)
df_prices$clean_prices <- round(df_prices$clean_prices, 2)
final_df <- left_join(final_df, df_prices, by = "id")


final_df$show <- gsub("-", " ", final_df$show)
final_df$theater_name <- gsub("-", " ", final_df$theater_name)

current_date <- as.Date("2024-03-22", format = "%Y-%m-%d")
current_date <- rep(current_date, times = 100)
df_current_date <- data.frame(current_date)
df_current_date$id <- c(1:100)

final_df <- left_join(final_df, df_current_date, by = "id")
final_df <- final_df %>% 
  select(id, everything())

final_df <- final_df %>% 
  rename(location = clean_location, prices = clean_prices)

final_df <- final_df %>% 
  relocate(dates, .after = prices)

final_df <- final_df %>% 
  mutate(days_elapsed = difftime(current_date, dates))

final_df <- left_join(final_df, df_1, by  = "id")


final_df$days_elapsed <- as.numeric(final_df$days_elapsed, units = "days")



#write.csv(final_df, "final_df.csv", row.names = FALSE) # I have saved df_final as a database in case Tomaticket removes any links that might prevent us from retrieving that observation again

final_df <- read_csv("final_df.csv")
#I remove observations 93 and 98 because the theater does not correspond to the show (in Tomaticket)
final_df <- final_df %>%
  slice(-93, -98)

final_df

```

## Descriptive Analysis

1.  **National Level**

```{r, include=TRUE, warning=FALSE}
ggplot(data = final_df, aes(x = reorder(location, location, length))) +
  geom_bar(aes(fill = location)) +
  geom_text(stat='count', aes(label=..count..), vjust=-0.5) +
  labs(title = "Number of shows per city",
       x = "City",
       y = "Number of shows") +
  theme_minimal()
```

At Tomaticket, there are 98 theater productions available, of which 82 are in theaters in Madrid, 13 in Barcelona, and one in Granada, Málaga, and Valladolid.

```{r}
ggplot(final_df, aes(x = days_elapsed)) +
  geom_density(fill = "skyblue", color = "navyblue") +
  labs(title = "Distribution of duration (available days)", x = "Days", y = "Density") +
  theme_minimal()

#Mean and mode
media <- mean(final_df$days_elapsed)
moda <- names(sort(-table(final_df$days_elapsed)))[1]
media
moda
```

We can observe that a large part of the theater productions are available in theaters for less than 250 days. On average, the shows advertised on Tomaticket are available in mean in each theater for about 256 days; however, the usual or typical availability is around 182 days (6 months).

```{r}
avg_price_per_city <- final_df %>%
  group_by(location) %>%
  summarise(avg_price = mean(prices, na.rm = TRUE))

ggplot(avg_price_per_city, aes(x = reorder(location, avg_price), y = avg_price, fill = avg_price)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("€%.2f", avg_price)), vjust = -0.5, color = "black", size = 3) +  # Añadir etiquetas de texto con los precios
  labs(title = "Average Price by Location",
       x = "Location",
       y = "Average price") +
  scale_fill_gradient(low = "lightgreen", high = "orange") +  # Escala de color de menos a más precio
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

Valladolid has the highest average price (€22) for the theater productions available on Tomaticket, followed by Barcelona (€15.53) and Madrid (€14.15). On the other hand, the cities with the cheapest theater productions are Granada and Málaga, with prices below €10.50.

However, for Valladolid, Granada, and Málaga, we only have one theater production available. This means that the average price for those cities will be equal to the price of that single observation.

In practical terms, this may result in a biased representation of the average price for those cities, as the price of a single observation may not be representative of the price distribution in that particular city. Therefore, it is important to consider this bias when interpreting the results of the analysis.

Therefore, we will focus our analysis on the theater productions advertised on Tomaticket only for Madrid, and we will compare between theaters.

2.  **Analyzing Madrid´s theaters**

```{r}
data_madrid <- final_df %>% 
  filter(location == "Madrid")
```

```{r}

precio_medio_por_teatro <- data_madrid %>%
  group_by(theater_name) %>%
  summarize(precio_medio = mean(prices))




ggplot(precio_medio_por_teatro, aes(x = reorder(theater_name, precio_medio), y = precio_medio)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Average Price per Theater",
       x = "Theater",
       y = "Average Price") +
  theme(axis.text.x = element_text(angle = 40, size = 6, hjust = 1))

```

Let´s see the most expensive and the cheapest theaters in Madrid

```{r}
data_madrid_tp <- data_madrid %>%
  group_by(theater_name) %>%
  summarise(avg_price = round(mean(prices, na.rm = TRUE), 2)) # 27 theaters
```

Most expensive theaters:

```{r}
data_madrid_tp %>% slice_max(avg_price, n = 6)
```

Cheapest theaters:

```{r}
data_madrid_tp %>% slice_min(avg_price, n = 6)
```

```{r}
obras_por_teatro <- data_madrid %>%
  group_by(theater_name) %>%
  summarize(num_obras = n())


ggplot(obras_por_teatro, aes(x = reorder(theater_name, num_obras), y = num_obras, fill = num_obras)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") + 
  labs(title = "Número de Obras por Teatro",
       x = "Teatro",
       y = "Número de Obras") +
  coord_flip() +
  geom_text(aes(label = num_obras), hjust = -0.2, size = 3)  
```

"Teatro Luchana" is the theater that hosts the most shows in Madrid (17), followed by "Teatro Lara" (10), "Teatro Arlequín" (9), and "Teatro Off Latina" (7).

```{r}
largest_shows_madrid <- data_madrid %>%
  top_n(5, days_elapsed)
largest_shows_madrid


ggplot(data = largest_shows_madrid,
                  aes(x = reorder(show, -days_elapsed),
                      y = days_elapsed)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Show", y = "Duración (días desde su fecha hasta ahora)", title = "Duración de los 5 shows más largos") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The show 'Doble o Nada' stands out significantly above the rest, with a run time of over 1000 days. The other shows closely following it have an approximate duration of 500 days each.

```{r}


plot_ly(data = data_madrid, x = ~days_elapsed, y = ~prices, text = ~paste("Nombre del show: ", show, "<br>Duración: ", current_date - dates, " días<br>Precio: $", prices), hoverinfo = "text",
        type = 'scatter', mode = 'markers', marker = list(color = ~prices, colorscale = 'RdYlGn',
                                                         colorbar = list(title = "Precio", ticksuffix = "$"))) %>%
  layout(title = "Relación entre Días Activos del Show y Precio Medio",
         xaxis = list(title = "Días Activos del Show"),
         yaxis = list(title = "Precio Medio"))
```

In shows priced at €10 or higher, there is a certain positive relationship observed between the number of days the show is active and its price. Although it is not a very strong relationship.

```{r}
teatros_con_mas_shows <- data_madrid %>%
  group_by(theater_name) %>%
  summarize(num_obras = n(),
            prices = mean(prices),
            days_elapsed = mean(days_elapsed)) %>%
  arrange(desc(num_obras))
```

```{r}
data_clustering <- teatros_con_mas_shows  %>%
  select(theater_name, prices, num_obras, days_elapsed)

#normalization
data_clustering_norm <- data_clustering %>%
  select(-theater_name) %>%
  scale()

#K.means clustering
set.seed(123)
kmeans_model <- kmeans(data_clustering_norm, centers = 3)
teatros_con_mas_shows$cluster <- as.factor(kmeans_model$cluster)



```

```{r}

grafico <- ggplot(teatros_con_mas_shows, aes(x = prices, y = days_elapsed, color = cluster, 
                                             text = paste("Teatro:", theater_name, "<br>Precio Medio:", prices, "<br>Days Elapsed:", days_elapsed))) +
  geom_point() +
  stat_ellipse(type = "norm", level = 0.95, linetype = "dashed") +
  labs(title = "Clustering de Teatros",
       x = "Precio Medio",
       y = "Days elapsed") +
  scale_color_manual(values = c("blue", "green", "purple"))


grafico_interactivo <- ggplotly(grafico, tooltip = "text") %>%
  plotly::layout(hovermode = "closest")


grafico_interactivo

```

We see that we could group the 27 theaters offering shows on Tomaticket into three different groups, taking into account their price and the number of days they are on the billboard.

Group 1, formed by those theaters whose offered shows have a relatively high average price and are also available on the billboard for more than 350 days.

Group 2, formed by theaters with the lowest prices and whose shows, on average, are available on the billboard for less than 300 days.

And finally, Group 3, which consists of theaters where the availability of shows is less than 300 days, but their prices are relatively high.

```{r, include=TRUE, warning=FALSE}


data_madrid <- data_madrid %>%
  mutate(Longitud = as.numeric(Longitud),
         Latitud = as.numeric(Latitud)) %>%
  na.omit()


map <- leaflet(data_madrid) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = median(data_madrid$Longitud), lat = median(data_madrid$Latitud), zoom = 9)

color_pal <- colorFactor(
  palette = c( "green", "red"), 
  domain = data_madrid$prices  
)

map <- map %>% 
  addCircles(lng = ~Longitud, 
             lat = ~Latitud,
             color = ~color_pal(prices),
              popup = ~paste("Teatro:",theater_name, "<br>Precio:" ,prices))


map <- map %>% 
  addLegend(position = "bottomleft", pal = color_pal, values = ~prices, bins = 4) 
  



map

```

Regarding the location of theaters in Madrid whose shows are offered on Tomaticket, the most notable aspect is that, at least for those that Tomaticket provides location information for, all the theaters are located in the city center.

If we consider the relationship between location and price, we do not observe any notable pattern.
